Commutative character sheaves and geometric types for supercuspidal representations

### 2017-05-02 Geometric types.tex
Added new proof of the theorem

\begin{lemma}
If $(\cs{L},\mu,\beta)\in \CS\ab(G)$ then $\mu  : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ is a morphism in $\CS\ab(G\times G)$.
\end{lemma}
\begin{proof}
Set $G^2 = G\times G$. Define $m^2 : G^2\times G^2 \to G^2$ by $m^2(g_1,g_2,g_1'g_2') = (g_1g_1',g_2g_2')$. Also define $p^2_i : G^2\times G^2 \to G^2$ by $p^2_i(g_1,g_2,g_1'g_2') = g_i g_i'$.
First we show that $m^*\cs{L}$ is an object in $\CS(G^2)$.
If $(\cs{L},\mu) \in \CS(G)$ then $m^*\cs{L}$ equipped with the isomorphism $(m^2)^* (m^*\cs{L}) \to m^*\cs{L} \boxtimes m^*\cs{L}$ defined by the diagram below.
\[
\begin{tikzcd}
(m^2)^* (m^*\cs{L}) \arrow{d}{(m^2)^*\mu} \arrow{rr} && m^*\cs{L} \boxtimes m^*\cs{L}\\
 (m^2)^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & (m^2)^*(p_1)^*\cs{L} \otimes (m^2)^*(p_2)^*\cs{L} \arrow[equal]{r} & (p^2_1)^*m^*\cs{L} \otimes (p^2_2)^*m^*\cs{L} \arrow[equal]{u}
\end{tikzcd}
\] 
\todo{Clifton is working here}
\end{proof}

@@

\begin{theorem}\label{thm:Gab}
%Let $G$ be a smooth group scheme over a finite field $\Fq$. Let $\CCS(G)$ be the category of {\it commutative characters sheaves} on $G$; see Definition~\ref{def:CCS}.
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories \todo{Is this trivial now?}
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}

\begin{proof}
\[
\begin{tikzcd}
{} & \arrow{dl}[swap]{q^*\ab} \CS(G\ab)\\
\CS\ab(G) & \arrow{l} \CS\ab(G\ab) \arrow{u}[swap]{\text{equiv.}}\\
\CCS(G) \arrow[>->]{u} & \arrow{l} \CCS(G\ab) \arrow[>->]{u}\\
\end{tikzcd}
\]
\todo{Clifton is working here}
\[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}
 &&&  m\ab^*\cs{L}\ab \arrow{dd}[swap]{\xi} \arrow{r}{\mu\ab} 
   & \cs{L}\ab\boxtimes\cs{L}\ab \arrow{dd}{\vartheta}\\
%
 && \  &\arrow{l}[swap]{(q^2)^*\ab}  & & &\\ 
%
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L}) 
 &&&   \theta^*(m\ab^*\cs{L}\ab) \arrow{r}{{\theta}^*\mu\ab} 
 &  \theta^*(\cs{L}\ab\boxtimes\cs{L}\ab)
 \end{tikzcd}
\]

By Definition~\ref{def:CCScom}, $\CCS(G\ab) \to \CS(G\ab)$ is a full subcategory.
Since the comparison functor $\CS(G\ab) \to \CS_{G\der}(G)$ is an equivalence, it determines a full subcategory $\CCS_{G\der}(G)$ which is equivalent to $\CCS(G\ab)$, as pictured below.
The proof of Theorem~\ref{thm:Gab} now reduces to the following claim: the essential image of $\CCS_{G\der}(G)$ under the equivalence $\CS_{G\der}(G) \to \CSab(G)$ is $\CCS(G)$.
\[
\begin{tikzcd}[row sep=30, column sep=40]
\CSab(G) & \CS(G\ab) \arrow{l}{\text{equiv}}[swap]{q\ab^*} \\
\CCS(G) \arrow{u}{\text{full sub}} & \CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{l}[swap]{\text{equiv}}
\end{tikzcd}
%\begin{tikzcd}[row sep=30, column sep=40]
%\CS(H) 
%& \arrow{l}[swap]{j^*} \CS(G) 
%&\arrow{l}[swap]{q^*} 
%\CS(G\ab) \arrow{dl}{\text{equiv}}[swap]{q_{H}^*} \\
%\CS'(G)  \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}} 
%&  \CS_{H}(G) \arrow{u}{\text{forget}}  
%&  %\arrow[bend left=35pt, dashed]{dll} 
%\CCS(G\ab) \arrow{u}[swap]{\text{full sub}} \arrow{dl}{\text{equiv}} \\
%\CCS(G) \arrow{u}{\text{full sub}} \arrow{r}{\beta \mapsto \nu}[swap]{\text{equiv}}
%& \arrow{u}{\text{full sub}}  \CCS_{H}(G) 
%& 
%\end{tikzcd}
\]

Let $(\cs{L},\mu,\nu)$ be the image of $(\cs{L},\mu,\beta)\in \CSab(G)$ and of $(\cs{L}\ab,\mu\ab)$ under $q\ab^*$;
we must show that $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}\ab$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{H} \to (\EE)_{H}$ as in Section~\ref{ssec:noncomdef}.
Recall the equivalence $L_2 : \Loc(G^2\ab) \to  \Loc_{G^2\der}(G^2)$ from Proposition~\ref{prop:CSGabab}.
The diagram in Definition~\ref{def:CCS} is precisely the result of applying the functor $L_2$ to the diagram in Definition~\ref{def:CCScom}, as pictured below; 
in particular $\gamma = (q^2)^* \xi$.
  \[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m_0^*\cs{L}_0 \arrow{dd}[swap]{\xi} \arrow{r}{\mu_0} 
   & \cs{L}_0\boxtimes\cs{L}_0 \arrow{dd}{\vartheta}
&&&   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}\\
 && \arrow{r}{(q^2)^*}  &\ & & &\\ 
   \theta^*(m_0^*\cs{L}_0) \arrow{r}{{\theta}^*\mu_0} 
 &  \theta^*(\cs{L}_0\boxtimes\cs{L}_0)
&&&   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L})
  \end{tikzcd}
\]
The diagram on the right is a diagram in $\Loc_{G^2\der}(G^2)$.
Using the equivalence $L_2$, this corresponds to a diagram in $\Loc(G^2\ab)$, necessarily the diagram on the left, above.
It follows that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$
if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G)$.
\end{proof}

>>>>

\begin{lemma}\label{lemma:morinCSab}
If $(\cs{L},\mu,\beta)\in \CS\ab(G)$ then $\mu  : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$, $\gamma : m^*\cs{L} \to  \theta^*(m^*\cs{L})$ and $\vartheta: \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes \cs{L})$ are morphisms in $\CS\ab(G\times G)$.
\end{lemma}
\begin{proof}
Set $G^2 = G\times G$. Define $m^2 : G^2\times G^2 \to G^2$ by $m^2(g_1,g_2,g_1'g_2') = (g_1g_1',g_2g_2')$. Also define $p^2_i : G^2\times G^2 \to G^2$ by $p^2_i(g_1,g_2,g_1'g_2') = g_i g_i'$.
First we show that $m^*\cs{L}$ is an object in $\CS(G^2)$.
If $(\cs{L},\mu) \in \CS(G)$ then $m^*\cs{L}$ equipped with the isomorphism $\mu^2: (m^2)^* (m^*\cs{L}) \to m^*\cs{L} \boxtimes m^*\cs{L}$ defined by the diagram below.
\[
\begin{tikzcd}
(m^2)^* (m^*\cs{L}) \arrow{d}{(m^2)^*\mu} \arrow{rr}{\mu^2} && m^*\cs{L} \boxtimes m^*\cs{L}\\
 (m^2)^*(\cs{L}\boxtimes\cs{L}) \arrow[equal]{r} & (m^2)^*(p_1)^*\cs{L} \otimes (m^2)^*(p_2)^*\cs{L} \arrow{r} & (p^2_1)^*m^*\cs{L} \otimes (p^2_2)^*m^*\cs{L} \arrow[equal]{u}
\end{tikzcd}
\] 
\todo{Clifton is working here}
\end{proof}

@@

\begin{theorem}\label{thm:Gab}
%Let $G$ be a smooth group scheme over a finite field $\Fq$. Let $\CCS(G)$ be the category of {\it commutative characters sheaves} on $G$; see Definition~\ref{def:CCS}.
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}

\begin{proof}
By definition, $\CCS(G)$ is a full subcategory of $\CS\ab(G)$; likewise, $\CCS(G\ab)$ is a full subcategory of $\CS\ab(G\ab)$.
We have just seen that $\CS\ab(G\ab)$ is equivalent to $\CS(G\ab)$.
By Proposition~\ref{prop:CSGabab}, pullback along the abelianization $q : G \to G\ab$ induces an equivalence $q^*\ab : \CS\ab(G)\to CS\ab(G)$.
Thus, the functor $\CS\ab(G\ab)\to \CS\ab(G)$ induced by pullback along $q$ is an equivalence.
The functor $\CCS(G\ab)\to \CCS(G)$ under consideration is the restriction of $\CS\ab(G\ab)\to \CS\ab(G)$ to the subcategory $\CCS(G\ab)$. 
\[
\begin{tikzcd}
{} & \arrow{dl}[swap]{q^*\ab} \CS(G\ab)\\
\CS\ab(G) & \arrow{l} \CS\ab(G\ab) \arrow{u}[swap]{\text{equiv.}}\\
\CCS(G) \arrow[>->]{u} & \arrow{l} \CCS(G\ab) \arrow[>->]{u}
\end{tikzcd}
\]
To prove the theorem, it is now sufficient to show that $\CCS(G\ab)\to \CCS(G)$ is essentially surjective.
Suppose $(\cs{L},\nu,\beta)\in \CCS(G)$.
Then $(\cs{L},\nu,\beta)\in \CS\ab(G)$.
Let $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ be given by the equivalences above.
Let $\xi : m\ab^*\cs{L}\ab \to \theta^* m\ab^*\cs{L}\ab$ be the isomorphism attached to $(\cs{L}\ab,\mu\ab)\in \CS(G\ab)$ as in Section~\ref{sec:comcom}.
Let $\gamma : m^*\cs{L} \to \theta^* m^*\cs{L}$ be the isomorphism attached to $\beta : \cs{L}\vert_{G\der} \to (\EE)_{G\der}$ as in Section~\ref{ssec:noncomdef}.
By Lemma~\ref{lemma:morinCSab}, the diagram on the right, below, is in $\CS(G\ab)$ and the diagram on the left, below, is in $\CS\ab(G)$.
\[
  \begin{tikzcd}[row sep=10, column sep = 20]
   m^*\cs{L} \arrow{dd}[swap]{\gamma} \arrow{r}{\mu} 
 & \cs{L}\boxtimes\cs{L} \arrow{dd}{\vartheta}
 &&&  m\ab^*\cs{L}\ab \arrow{dd}[swap]{\xi} \arrow{r}{\mu\ab} 
   & \cs{L}\ab\boxtimes\cs{L}\ab \arrow{dd}{\vartheta}\\
%
 && \  &\arrow{l}[swap]{(q^2)^*\ab}  & & &\\ 
%
   \theta^*(m^*\cs{L}) \arrow{r}{{\theta}^*\mu} 
&  \theta^*(\cs{L}\boxtimes\cs{L}) 
 &&&   \theta^*(m\ab^*\cs{L}\ab) \arrow{r}{{\theta}^*\mu\ab} 
 &  \theta^*(\cs{L}\ab\boxtimes\cs{L}\ab)
 \end{tikzcd}
\]
The diagram on the left, above, is precisely the result of applying the functor $(q^2)^*\ab$ to the diagram on the right, above; in particular $\gamma = (q^2)^*\ab \xi$.
Since $(q^2)^*\ab$ is an equivalence by Proposition~\ref{prop:CSGabab}, it follows that the diagram in Definition~\ref{def:CCS} commutes if and only if the diagram in Definition~\ref{def:CCScom} commutes. 
In other words, $(\cs{L},\mu,\beta)\in \CCS(G)$ if and only if $(\cs{L}\ab,\mu\ab)\in \CCS(G\ab)$.

\end{proof}
